---
title: "Assignment 11"
author: "Gina Catellier"
date: "10/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

1. Install the package `mlbench` and use the follows to import the data

```{r}
library(mlbench)
data(PimaIndiansDiabetes)
df <- PimaIndiansDiabetes
```

- Set seed to be 2020. 
- Partition the data into 80% training and 20% testing.  

```{r}
library(caret)
set.seed(2020)

names(df)[9] <- 'target'
splitIndex <- createDataPartition(df$target, p = .80,
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

-------

2. Tuning Decision Tree:  Use cross-validation with 10 k-folds to find the maxdepth with the greatest accuracy. Plot the accuracy associated with different maxdepths against the maxdepths. The range to search for maxdepth is from 1 to 10. 

```{r}
tuneGrid = expand.grid(maxdepth = 1:10)

trControl = trainControl(method = "cv",
                         number = 10)

tree_approach1 <- train(target~., data=df_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach1)
```

-------

3. Make the final decision to select the maxdepth for your decision tree.  Is your selected maxdepth the same as the maxdepth found in 2. 

-------

The maxdepth producing the greatest accuracy is maxdepth 1, which is also the lowest mapdepth producing top accuracy. This selection is the same maxdepth as found in 2.

-------

4. Calculate the accuracy of your decision tree (the decision tree with your selected maxdepth in 3) on the test data. 

```{r}
pred <- predict(tree_approach1, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = "pos")
cm$overall[1]
```

-------

5. Redo 2-4 with an alternative method to cross-validation. 

```{r}
tuneGrid = expand.grid(maxdepth = 1:10)

trControl = trainControl(method = "LGOCV",
                         number = 10)

tree_approach2 <- train(target~., data=df_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)

plot(tree_approach2)
```


This approach would require the selection of a maxdepth of 8, which is the maxdepth that provides the greatest accuracy.

```{r}
pred2 <- predict(tree_approach2, df_test)
cm2 <- confusionMatrix(data = pred2, reference = df_test$target, positive = "pos")
cm2$overall[1]
```

