---
title: "Assignment 6"
author: "Gina Catellier"
date: "9/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**Questions**

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides.

```{r}
installed.packages('tidyverse')
library(tidyverse)
df <- read_csv('titanic.csv')
```

-------

1. For one continuous variable:

- Density

```{r}
df %>% ggplot() + geom_density(mapping = aes(x=Age))
```

This graph shows that a majority of the passengers on the titanic were around 25 years old, as this spot has the highest density of around 0.03.

- Histogram 

```{r}
df %>% ggplot() + geom_histogram(mapping = aes(x=Age))
```

This graph shows that the most common ages of passengers on the titanic were around 18 to 25 years old, as these bars have the highest count in the histogram.

- Boxplot

```{r}
df %>% ggplot() + geom_boxplot(mapping = aes(x=Age))
```

This graph shows the middle 50% of passengers on the titanic were between 20 and 38 years old. There are also passengers that were outliers, with ages over 65.

-------

2. For one categorical variable

- Barplot

```{r}
df %>% ggplot() + geom_bar(mapping = aes(x=Survived))
```

This graph shows that the count of passengers that did not survive exceeds the count of passengers that did survive.

-------

3. For one two continuous variables

- Scatter Plot

```{r}
df %>% ggplot() + geom_point(mapping = aes(x=Age, y=Fare))
```

This graph shows that there is seemingly no correlation between the age of a passenger and the price that they paid on the ticket fare.

- Line Plot

```{r}
df %>% ggplot() + geom_line(mapping = aes(x=Age, y=Fare))
```

This graph shows that there is little correlation between the age of a passenger on the titanic and the price that they paid for the ticket fare. However, there does seem to be a spike around the age of the late 30s.

- Smooth Plot

```{r}
df %>% ggplot() + geom_smooth(mapping = aes(x=Age, y=Fare))
```

This graph shows that the lowest fare price for titanic passengers was for ages around 25 and the highest was for ages around 55.

-------

4. For one continuous + one categorical variables

- Density

```{r}
df %>% ggplot() + geom_density(mapping = aes(x=Age, color = Sex))
```

This graph shows that the majority of passengers on the titanic were around 25 years old, for both men and women.

- BoxPlot

```{r}
df %>% ggplot() + geom_boxplot(mapping = aes(x=Age, fill = Sex))
```

This graph shows that, although similar, the average age for a male passenger on the titanic is slightly higher than that for a female passenger.

-------

5. For two categorical variables: barplot
```{r}
df %>% ggplot() + geom_bar(mapping = aes(x=Survived, fill=Sex))
```

This graph shows that of the larger amount of passengers that did not survive the titanic, a majority of them were male. On the other hand, of the smaller portion of passengers that did survive most of them were female.
