---
title: "Assignment 13"
author: "Gina Catellier"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

#### NOTICE: Question 7 and 8 are optional. 

-------

1. Write a function to calculate the age of a person.  Give an example to test your function. 

- Input: a number, x, (year born)

- Output: Print out "You are r age". Where r is the age of the person, i.e. 2020 - x. 

Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#29)

```{r}
age <- function(x)
{
  r <- 2020 - x
  print(paste0('You are ', r, ' age.'))
}
```

```{r}
age(2000)
```

-------

2. Write the following function. Give an example to test your function. 

- Input: a number

- Output: print out: "You input an even number!" if the number is event, or "You input an odd number!" otherwise. 

Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#33)

```{r}
odd_even <- function(x)
{
  if(x %% 2==0)
  {
    print('You input an even number!')
  }
  else
  {
    print('You input an odd number!')
  }
}
```

```{r}
odd_even(3)
```

-------

3. Write the following function.  Give an example to test your function. 

- Input: a numeric vector 

- Output:  

  - if the input vector has missing values: return the input vector with missing values replaced by mean
  
  - if the input vector has no missing value: return the same input vector
  
Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#37)

```{r}
mean_impute <- function(x)
{
    mean_of_x <- (mean(x, na.rm=TRUE))
    library(tidyr)
    x <- replace_na(x, mean_of_x) 
    return(x)    
}
```

```{r}
library(tidyverse)
df <- read_csv('titanic.csv')
x1 <- mean_impute(df$Age)
sum(is.na(x1))
```

-------

4. Write the following function.  Give an example to test your function. 

- Input: a vector x 

- Output:  The vector x where the missing values replaced by the mean (if x is numeric) or the mode (if x is non-numeric).  If x does not have missing value, return the same vector x.  

Hint: Combine the function in Question 3 and [this function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#37)

```{r}
mean_mode_impute <- function(x)
{
  if(!is.numeric(x))
  {
    mode_of_x <- names(sort(-table(x)))[1]
    library(tidyr)
    x <- replace_na(x, mode_of_x) 
  }
  else
  { 
    mean_of_x <- (mean(x, na.rm=TRUE))
    library(tidyr)
    x <- replace_na(x, mean_of_x) 
    return(x)
  }
}
```

```{r}
x2 <- mean_mode_impute(df$Embarked)
sum(is.na(x2))
```

-------

5. Write the following function. Give examples to test your function. 

- Input: A data frame of two variables x and y

- Output: 

 - A boxplot of x by y (`aes(x = x, fill = y)`) if x is numeric and y is non-numeric
 
 - A boxplot of y by x (`aes(x = x, fill = y)`) if y is numeric and x is non-numeric
 
 - print out 'This function cannot visualize your data' otherwise

Hint: 

- [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#41)
 
- You can refer to this slide to plot a boxplot: https://bryantstats.github.io/math421/slides/6_viz.html#36

```{r}
boxplot_function <- function(x)
{
  if(is.numeric(x[[1]])&(!(is.numeric(x[[2]]))))
  {
    x %>% ggplot(aes(x = x[[1]], fill = x[[2]]))+
      geom_boxplot()+
      labs(x = names(x)[1], fill = names(x)[2])
  }
  else if (!(is.numeric(x[[1]])|is.numeric(x[[2]])))
  {
    x %>% ggplot(aes(x = x[[2]], fill = x[[1]]))+
      geom_boxplot()+
      labs(x = names(x)[2], fill = names(x)[1])
  }
  else 
  {
    print('This function cannot visualize your data.')
  }
}
```

```{r}
example <- df %>% select(Age, Sex)
boxplot_function(example)
```

-------
 
6. Combine the function in Question 5 and the function in this example ([Link]()) to have a function that can handle all possible cases. 

```{r}
boxplot_scatter_function <- function(x)
{
  if(is.numeric(x[[1]])&(!(is.numeric(x[[2]]))))
  {
    x %>% ggplot(aes(x = x[[1]], fill = x[[2]]))+
      geom_boxplot()+
      labs(x = names(x)[1], fill = names(x)[2])
  }
  else if (!(is.numeric(x[[1]])|is.numeric(x[[2]])))
  {
    x %>% ggplot(aes(x = x[[2]], fill = x[[1]]))+
      geom_boxplot()+
      labs(x = names(x)[2], fill = names(x)[1])
  }
  else if(is.numeric(x[[1]])&is.numeric(x[[2]]))
  {
    x %>% ggplot(aes(x = x[[1]], y = x[[2]]))+
      geom_point()+
      labs(x = names(x)[1], y = names(x)[2])
  }
  else if (!(is.numeric(x[[1]])|is.numeric(x[[2]])))
  {
    x %>% ggplot(aes(x = d[[1]], fill = x[[2]]))+
      geom_bar(position = 'dodge')+
      labs(x = names(x)[1], fill = names(x)[2])
  }
  else 
  {
    print('This function cannot visualize your data.')
  }
}
```

```{r}
example2 <- df %>% select(Age, Fare)
boxplot_scatter_function(example2)
```

-------

7. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 
  - train_percent: a number presenting a proportion of training data. The default train_percent is .8

- Output: the accuracy of the decision model `rpart` where the training data is train_percent. 

```{r}
rpart_model <- function(input_data, train_percent = 0.8)
{
library(caret)
set.seed(00000)
splitIndex <- createDataPartition(input_data$target, p = train_percent, list = FALSE)
df_train <- input_data[ splitIndex,]
df_test <- input_data[-splitIndex,]
library(rpart)
tree1<-rpart(target ~ ., data = df_train)

library(rattle)
fancyRpartPlot(tree1)
}
```

```{r}
df2 <- df
names(df2)[2] <- 'target'

df2$PassengerId =  NULL
df2$Ticket =  NULL
df2$Name = NULL
df2$Cabin = NULL

df2$target <- factor(df2$target)
df2$Pclass = factor(df2$Pclass)
df2$Sex <- factor(df2$Sex)
df2$Embarked <- factor(df2$Embarked)
df2$Parch <- factor(df2$Parch)

df2$Age[is.na(df2$Age)] = mean(df2$Age, na.rm = TRUE)
df2 = drop_na(df2)

rpart_model(df2, 0.8)
```

-------

8. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 
  
  - train_percent: a number presenting a proportion of training data. The default train_percent is .8

- Output: the plot of variable important by random forest trained by caret. 

```{r}
random_forest_model <- function(input_data, train_percent=.8)
{
library(caret)
set.seed(00000)
splitIndex <- createDataPartition(input_data$target, p = train_percent, list = FALSE)
df_train <- input_data[ splitIndex,]
df_test <- input_data[-splitIndex,]
library(randomForest)
model = randomForest(target ~ ., data=df_train, ntree = 1000)

plot(importance(model))
}
```

```{r}
random_forest_model(df2, 0.8)
```

