---
title: "Assignment 3"
author: "Gina Catellier"
date: "9/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Problems
-------

1. Use `read.csv` to import the Covid19 data from WHO: [link](https://covid19.who.int/WHO-COVID-19-global-data.csv)

```{r}
df <- read.csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
```

2. How many columns and rows does the data have? How many categorical variables does the data have?

```{r}
dim(df)
table(sapply(df, class))
```


3. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?

```{r}
sum(is.na(df))
colSums(is.na(df))
which.max(colSums(is.na(df)))

```


4. Correct the types of the variables that are mismatched type.

```{r}
str(df)
 # install.packages('lubridate')
library(lubridate)
df$ï..Date_reported <- ymd(df$ï..Date_reported)
class(df$ï..Date_reported)

```


5. Find the daily average number of cases globally.

```{r}
mean(df$New_cases)
```


6. Find the maximum cases a day.  

```{r}
max(df$New_cases)
```


7. Which country at which date has the maximum number of cases a day? *Hint:  use the `which.max` function to find the index of the row that contains the maximum number of cases. Then from the row index, find that country and the date. * 

```{r}
r = which.max(df$New_cases)
df$Country[r]
df$ï..Date_reported[r]
```


8.  How many WHO's region there are in the data?

```{r}
dim(table(df$WHO_region))
```

9.  Find the daily average cases in EUROPE. Find the daily average cases by regions.

```{r}
dfEurope <- df[df$WHO_region=='EURO',]
mean(dfEurope$New_cases)
```


10. Find the daily average number of cases in the US in the last 5 days. 

```{r}
df2 = df[ (df$Country=='United States of America') & (Sys.Date()- df$ï..Date_reported<5),]
mean(df2$New_cases)
```


11. Which day of the week has the most number of cases in South East Asia (SEARO region)?

```{r}
df3 <- df[df$WHO_region=='SEARO',]
df3$dayofweek <- wday(df3$ï..Date_reported)
table(df3$dayofweek)
w = which.max(table(df3$dayofweek))
df3$dayofweek[w]
wday(df3$dayofweek[w], label=TRUE)
```

