---
title: "Assignment 9"
author: "Gina Catellier"
date: "10/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


-------

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

Hint: 

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths

```{r}
library(tidyverse)
library(ggplot2)
library(gganimate)
library(gifski)
df <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
```

```{r}
library(lubridate)
d1 <- df %>% filter(Date_reported == max(Date_reported))
d2 <- d1 %>% mutate(rank = rank(-Cumulative_deaths)) %>% ungroup()
d3 <- d2 %>% filter(rank <= 3)
```

```{r}
d3$Country
```

2. Make a plot revealing the number of deaths in the three countries with the most numbers of deaths

```{r}
df %>% 
  filter(Country %in% d3$Country) %>%
  group_by(Country) %>%
  ggplot(aes(y=Cumulative_deaths,
            x=Date_reported,
            color=Country,
            fill=Country))+ 
  labs(title = 'Cumulative Deaths over Time', caption
            = 'This graph shows the number of deaths 
            in the top 3 countries.', ylabel = 
              'Number of Deaths', xlabel = 'Date 
            Reported')+
  geom_point(size=3)+
  geom_text(aes(label = Cumulative_deaths), 
            hjust = -.1, size=3)+
  geom_line()+
  transition_reveal(Date_reported)
```


3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 

```{r}
df$deaths_per_cases = df$Cumulative_deaths/df$Cumulative_cases
a1 <- df %>% filter(Date_reported == max(Date_reported))
a2 <- a1 %>% mutate(rank = rank(-deaths_per_cases)) %>% ungroup()
a3 <- a2 %>% filter(rank <= 3)
```

```{r}
a3$Country
```

4. Make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 

```{r}
df %>% 
  filter(Country %in% c('United States of America', 'Italy', 'Mexico'))%>%
  group_by(Country) %>%
  ggplot(aes(y=deaths_per_cases,
            x=Date_reported,
            color=Country,
            fill=Country))+ 
  labs(title = 'Cumulative Deaths per Cases over 
       Time', caption = 'This graph shows the number 
       of deaths per cases in the top 3 countries.', 
       ylabel = 'Number of Deaths per Case', xlabel         ='Date Reported')+
  geom_point(size=3)+
  geom_text(aes(label = deaths_per_cases), 
            hjust = -.1, size=3)+
  geom_line()+
  transition_reveal(Date_reported)
```

