---
title: "Assignment 14"
author: "Gina Catellier"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

1. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: a data frame with all the missing of numeric variables replaced by the associated means.

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/14_writing_functions.html#28)

```{r}
mean_impute <- function(x)
{
    mean_of_x <- (mean(x, na.rm=TRUE))
    library(tidyr)
    x <- replace_na(x, mean_of_x) 
    return(x)    
}

numeric_impute <- function(x)
{
  for (i in 1:length(x))
  {
    x[[i]] <- mean_impute(x[[i]])
  }
  return(x)
}
```

```{r}
library(tidyverse)
df <- read_csv('titanic.csv')
colSums(is.na(df))
```

```{r}
df1 <- numeric_impute(df)
colSums(is.na(df1))
```

-------

2. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: a data frame with all the missing of variables replaced by the associated means (for numeric variables) or modes (for non-numeric variables).

Hint: Combine the function in Problem 1 and the function in [this example](https://bryantstats.github.io/math421/gganimate/14_writing_functions.html#28)

```{r}
mean_mode_impute <- function(x)
{
  if(!is.numeric(x))
  {
    mode_of_x <- names(sort(-table(x)))[1]
    library(tidyr)
    x <- replace_na(x, mode_of_x) 
  }
  else
  { 
    mean_of_x <- (mean(x, na.rm=TRUE))
    library(tidyr)
    x <- replace_na(x, mean_of_x) 
    return(x)
  }
}

impute_all <- function(x)
{
  for (i in 1:length(x))
  {
    x[[i]] <- mean_mode_impute(x[[i]])
  }
  return(x)
}
```

```{r}
df2 <- impute_all(df)
colSums(is.na(df2))
```

-------

3. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: Bar plots of all non-numeric variables

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/14_writing_functions.html#32)

```{r}
bar_plot <- function(x)
{
  library(ggplot2)
  for (i in 1:length(x))
  {
    if (!is.numeric(x[[i]]))
    {
      print(ggplot(x, aes(x = x[[i]]))+ 
              geom_bar()+
              labs(x = names(x)[i]))
    }
  }
}
```

```{r}
bar_plot(df2)
```

-------

4. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: all possible the bar plots of a non-numeric variable filled by a non-numeric variable.

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/14_writing_functions.html#40)

```{r}
bar_plot2 <- function(x)
{
  library(ggplot2)
  l <- length(x)
  for (i in 1:(l-1))
    for (j in (i+1):l)
  {
    if (!is.numeric(x[[i]])& (!is.numeric(x[[j]])))
    {
      print(ggplot(x, aes(x = x[[i]], color = x[[j]]))+ 
              geom_bar()+labs(x = names(x)[i], color = names(x)[j]))
    }
  }
}
```

```{r}
bar_plot2(df2)
```

-------

5. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: 

  - all possible the bar plots of a non-numeric variable filled by a non-numeric variable.
  
  - all possible the density plots of a numeric variable colored by a non-numeric variable
  
  - all possible the scatter plots.

Hint: Combine [this function](https://bryantstats.github.io/math421/gganimate/14_writing_functions.html#36), [this function](https://bryantstats.github.io/math421/gganimate/14_writing_functions.html#39), and the function in Question 4. One way to combine is creating a new function, `quick_plot`, and call these three functions within `quick_plot`. 

```{r}
density_plot <- function(d)
{
  library(ggplot2)
  l <- length(d)
  for (i in 1:(l-1))
    for (j in (i+1):l)
  {
    if (is.numeric(d[[i]])& (!is.numeric(d[[j]])))
    {
      print(ggplot(d, aes(x = d[[i]], color = d[[j]]))+ 
              geom_density()+labs(x = names(d)[i], color = names(d)[j]))
    }
  }
}

scatter_plot <- function(d)
{
  library(ggplot2)
  l <- length(d)
  for (i in 1:(l-1))
    for (j in (i+1):l)
  {
    if (is.numeric(d[[i]])&is.numeric(d[[j]]))
    {
      print(ggplot(d, aes(x = d[[i]], y = d[[j]]))+ 
              geom_point()+
              labs(x = names(d)[i], y = names(d)[j]))
    }
  }
}

quick_plot <- function(x)
{
    bar_plot2(x)
    density_plot(x)
    scatter_plot(x)
}
  
```

```{r}
quick_plot(df2)
```

