---
title: "Midterm Presentation"
author: "Gina Catellier"
date: "10/22/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data Visualizations

```{r, message=FALSE}
library(tidyverse)
library(haven)
df = read_sas('hdd0318cy.sas7bdat')

df <- filter(df, yod=='18')

df <- select(df, yod, payfix, pay_ub92, age, sex, raceethn, provider, moa, yoa, mod, admtype, asource, preopday, los, service, icu, ccu, dispub92, payer, drg, trandb, randbg, randbs, orr, anes, seq, lab, dtest, ther, blood, phar, other, patcon, bwght, total, tot, ecodub92, b_wt, pt_state, diag_adm, ancilar, campus, er_fee, er_chrg, er_mode, obs_chrg, obs_hour, psycchrg, nicu_day)

df$payfix <- NULL
df$preopday <- NULL 
df$obs_hour <- NULL
df$nicu_day <- NULL

df$age2 <- case_when(
          df$age < 13 ~ 'child',
          df$age < 20 ~ 'teenager',
          TRUE ~ 'adult')
```
**Variables Focused On:**


- sex

- raceethn

- age & age2

- admtype

- los

- tot

## Price by Sex vs Price by Race

```{r, message=FALSE}
df %>% filter(sex == 1 | sex == 2) %>% group_by(sex) %>% summarise(average_price = mean(tot)) %>% ggplot() + geom_col(mapping=aes(x=sex, y=average_price, fill=sex)) + labs(title = 'Average Price by Sex', x = 'Sex', y='Average Price')
```

---

```{r, message=FALSE}
df %>% group_by(raceethn) %>% summarise(average_price = mean(tot)) %>% ggplot() + geom_col(mapping=aes(x=raceethn, y=average_price, fill=raceethn))+labs(title= 'Average Price by Race', x = 'Race', y = 'Average Price')
```

## Comparison

- Average price for male was about $5,000 higher than that for a female. 

- Average price for all races are around within $5,000 of each other.

- The highest average price was for race 1 (White) and the lowest average price was for race 3 (Asian).

## Correlation of Race vs Sex

```{r, message=FALSE}
df %>% filter(sex == 1 | sex == 2) %>% ggplot() + geom_bar(mapping=aes(x=raceethn, fill=sex), position='dodge') + labs(title = 'Count of Race by Sex', x = 'Race', y = 'Number of Patients')
```

---





**Correlation of Race vs Sex**


- A large majority of patients were of race 1 (white).

- All races had slightly more female patients than male patients.

- This connects to the fact that average cost for male is greater than average cost for female.


## Average Price by Age

```{r, message=FALSE}
df %>% group_by(age) %>% summarise(average_price = mean(tot)) %>% ggplot() + geom_col(mapping=aes(x=age, y=average_price, fill=age)) + labs(title = 'Average Price by Age', x = 'Age', y = 'Average Price')
```

---





**Average Price by Age**


- This shows a large spike in average price around the ages 10 to 20.

- The lowest average price was from ages 25 to 35.

- There is a bell curve for average price from ages 35 to 95.

## Total Cost by Age by Sex

```{r, message=FALSE}
df %>% filter(sex == 1 | sex == 2) %>% 
ggplot(mapping=aes(x=age, y=tot, color=sex))+ geom_point()+ facet_wrap(~sex) +labs(title = 'Total Cost by Age', y = 'Total Cost', x = 'Age')
```

---





**Total Cost by Age by Sex**


- This shows a very similar trend to that of average price by age.

- There is no significant difference between male and female for total cost by age.

## Total Cost by Age Categorical

```{r, message=FALSE}
df %>% group_by(age2) %>% summarise(average_price = mean(tot)) %>% ggplot() + geom_col(mapping=aes(x=age2, y=average_price, fill=age2)) +labs(title = 'Average Price by Age', x = 'Age', y = 'Average Price')
```

---





**Total Cost by Age Categorical**


- Average price for teenager is about $15,000 more than average price for child.

- Average price for teenager is about $5,000 more than average price for adult.

## Count of Admission Type by Sex

```{r, message=FALSE}
df %>% filter(sex == 1 | sex == 2) %>% 
ggplot(mapping=aes(x=admtype, fill=sex))+ geom_bar()+ facet_wrap(~sex) +labs(title = 'Count of Admission Type by Sex', x = 'Admission Type', y = 'Number of Patients')
```

---




**Count of Admission Type by Sex**

- Admission Type 1 (Emergency) is the most common admission type for both males and females.

- Females are slightly higher in every admission type than males.

## Count of Admission Type by Age

```{r, message=FALSE}
df %>% ggplot(mapping=aes(x=admtype, fill=age2))+ geom_bar()+ facet_wrap(~age2) + labs(title = 'Count of Admission Type by Age', x = 'Admission Type', y = 'Number of Patients')
```

---




**Count of Admission Type by Age**


- Adult very closely resembles the graph for both male and female.

- All age groups have Admission Types 1 and 2 (Emergency and Urgent) in their most popular types.

## Count of Admission Type by Age2

```{r, message=FALSE}
df %>% filter(age2 == 'child' | age2 == 'teenager') %>% ggplot(mapping=aes(x=admtype, fill=age2))+ geom_bar()+ facet_wrap(~age2) + labs(title = 'Count of Admission Type by Age', x = 'Admission Type', y = 'Number of Patients')
```


---




**Count of Admission Type by Age(Not Adult)**


- Now, both child and teenager results resemble that of an adult.

- The exception is the large amount of admission type 4 (newborn) for child.

## Average Length of Stay by Age

```{r, message=FALSE}
df %>% group_by(age2) %>% summarise(average_los = mean(los)) %>% ggplot(mapping=aes(x=age2, y=average_los, fill=age2))+ geom_col() + labs(title = 'Average Length of Stay by Age', x = 'Age', y = 'Average Length of Stay')
```


---




**Average Length of Stay by Age**



- For child and adult, the average length of stay is about 5 days.

- For teenager, the average length of stay is about 9 days.

- This could explain why teenager costs were so much higher, on average.

## Average Length of Stay by Age by Month of Admission

```{r step12811101, eval = FALSE, echo = TRUE}
library(gganimate)
library(gifski)

df %>% 
  group_by(moa,age2) %>%
  summarise(average_los = mean(los)) %>%
  ggplot(aes(y=average_los,
            x=moa,
            color=age2,
            fill=age2))+ labs(title = 'Average Length of Stay by Age by Month of Admission', x = 'Month of Admission', y = 'Average Length of Stay')+
  geom_point(size=3)+
  geom_text(aes(label = average_los), 
            hjust = -.1, size=3)+
  geom_line()+
  transition_reveal(moa)
```

---

```{r, ref.label = "step12811101", echo = FALSE, cache = TRUE, fig.height = 6, fig.width = 7}
```


## Predictive Modeling
**Target Variable = cost**

```{r, message=FALSE}
df$cost <- case_when(
          df$tot < mean(df$tot) ~ 'low_cost',
          TRUE ~ 'high_cost')

df1 <- select(df, age, sex, raceethn, provider, moa, mod, admtype, campus, los, cost)

df1 <- df1 %>% filter(raceethn != '', admtype != '')

library(mlbench)
library(caret)
set.seed(2020)

names(df1)[10] <- 'target'
splitIndex <- createDataPartition(df1$target, p = .90,
                                  list = FALSE)
df_train <- df1[splitIndex,]
df_test <- df1[-splitIndex,]
```

---

Decision Tree Using rpart

```{r, message=FALSE}
library(rpart)
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))

library(rattle)
fancyRpartPlot(tree_model)
```

---

Variable Importance

```{r, message=FALSE}
barplot(tree_model$variable.importance)
```


## Comparing Models

- rpart2

- blackboost

- ada

---

rpart2
```{r, message=FALSE}
tuneGrid = expand.grid(maxdepth = 1:5)

trControl = trainControl(method = "cv",
                         number = 5)

tree_approach1 <- train(target~., data=df_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach1)
```


---

blackboost
```{r, message=FALSE}
tuneGrid = expand.grid(maxdepth = 2:4, mstop=3)
tree_approach2 <- train(target~., data=df_train, 
                                method = "blackboost", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach2)
```


---

ada
```{r, include=FALSE}
tuneGrid = expand.grid(maxdepth = 2:4, iter=3, nu=2)

tree_approach3 <- train(target~., data=df_train, 
                                method = "ada", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
```

```{r}
plot(tree_approach3)
```

## Comparison

```{r, message=FALSE}
results <- resamples(list(rpart2 = tree_approach1,
                          blackboost = tree_approach2,
                          ada = tree_approach3))
bwplot(results)
```


## Accuracy of Final Model

pred <- predict(tree_approach3, df_test)
cm <- confusionMatrix(data = pred, reference = 
as.factor(df_test$target), positive = "high_cost")
cm$overall[1]

```{r, message=FALSE}
pred <- predict(tree_approach3, df_test)
cm <- confusionMatrix(data = pred, reference = 
as.factor(df_test$target), positive = "high_cost")
cm$overall[1]
```


## Target Variable = sex

```{r, message=FALSE}
df2 <- select(df, age, sex, raceethn, provider, moa, mod, admtype, campus, los, cost)

df2 <- df2 %>% filter(raceethn != '', admtype != '', sex == 1 | sex == 2)

names(df2)[2] <- 'target'
splitIndex <- createDataPartition(df2$target, p = .90,
                                  list = FALSE)
df2_train <- df2[splitIndex,]
df2_test <- df2[-splitIndex,]

```

rpart2
```{r, message=FALSE, echo=FALSE}
tuneGrid = expand.grid(maxdepth = 1:5)

trControl = trainControl(method = "cv",
                         number = 5)

tree_approach1_2 <- train(target~., data=df2_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach1_2)
```


---


blackboost
```{r, message=FALSE, echo=FALSE}
tuneGrid = expand.grid(maxdepth = 2:4, mstop=3)
tree_approach2_2 <- train(target~., data=df2_train, 
                                method = "blackboost", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach2_2)
```


---



ada
```{r, include=FALSE}
tuneGrid = expand.grid(maxdepth = 2:4, iter=3, nu=2)

tree_approach3_2 <- train(target~., data=df2_train, 
                                method = "ada", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)

```

```{r}
plot(tree_approach3_2)
```


##Comparison 2

```{r, message=FALSE}
results <- resamples(list(rpart2 = tree_approach1_2,
                          blackboost = tree_approach2_2,
                          ada = tree_approach3_2))
bwplot(results)
```

##Accuracy of Final Model 2

pred <- predict(tree_approach3_2, df2_test)
cm <- confusionMatrix(data = pred, reference = 
as.factor(df2_test$target), positive = "1")
cm$overall[1]

```{r, message=FALSE}
pred <- predict(tree_approach3_2, df2_test)
cm <- confusionMatrix(data = pred, reference = 
as.factor(df2_test$target), positive = "1")
cm$overall[1]
```

