---
title: "Midterm"
author: "Gina Catellier"
date: "10/14/2020"
output: html_document
---
-------

# Instruction

The midterm has two components: the Rmarkdown notebook (html) and the presentation.  The notebook and the presentation file are due Nov 26. We will do the presentation in class 

**The notebook:** The notebook should be created using `rmarkdown` (like other assignments). The notebook should have a title. 


**The Presentation:** Present your results in 5-10 minutes. To make the presentation using Rmarkdown, do the follows: 

    - In Rstudio -> File -> New File -> R markdown
    
    - In the left panel, click to Presentation -> Click OK
    
    - Now you have an Rmarkdown that can be knitted to be a html presentation 
    

**Data:**  The data for the mid-term project is the Rhode Island Department of Health Hospital Discharge Data.  Each row of the data presents a patient. 

Link: https://drive.google.com/open?id=15QNBf6YYKocK2nNIfpKDer58kQnCPNZJ 

-------

## I. Data Wranggling

1. Download the data file `hdd0318cy.sas7bdat`.  

2. Use `read_sas` in library `haven` to read the data.

```{r}
library(tidyverse)
library(haven)
df = read_sas('hdd0318cy.sas7bdat')
```
    
3. Filter the data to have only patients of the year 2018 (`yod==2018`)

```{r}
df <- filter(df, yod=='18')
```

    
4. Select to work with only following variables: 

```{r, eval=FALSE}
                      "yod", "payfix","pay_ub92","age",  
                      "sex","raceethn","provider","moa", 
                      "yoa","mod","admtype", "asource" , 
                      "preopday" ,"los", "service" , "icu","ccu",    
                      "dispub92", "payer"  ,"drg","trandb", 
                      "randbg","randbs","orr", "anes","seq",   
                      "lab","dtest", "ther","blood","phar", 
                      "other","patcon","bwght","total","tot" ,  
                      "ecodub92","b_wt","pt_state","diag_adm","ancilar" ,
                      "campus","er_fee","er_chrg","er_mode","obs_chrg",
                      "obs_hour","psycchrg","nicu_day"
```
 
```{r}
df <- select(df, yod, payfix, pay_ub92, age, sex, raceethn, provider, moa, yoa, mod, admtype, asource, preopday, los, service, icu, ccu, dispub92, payer, drg, trandb, randbg, randbs, orr, anes, seq, lab, dtest, ther, blood, phar, other, patcon, bwght, total, tot, ecodub92, b_wt, pt_state, diag_adm, ancilar, campus, er_fee, er_chrg, er_mode, obs_chrg, obs_hour, psycchrg, nicu_day)
```

 
*Notice*:  You may want to save the current data on your laptop if you want to read it again.  To save the data file use `write_csv(df, 'midterm.csv')`, for example.  

```{r}
write_csv(df, 'midterm.csv')
```

5. What are variables that have missing values?

```{r}
colSums(is.na(df))
```
 
6. Remove all variables with missing values

```{r}
df$payfix <- NULL
df$preopday <- NULL 
df$obs_hour <- NULL
df$nicu_day <- NULL
```
 
7. Refer to the data description in the file `HDD2015-18cy6-20-19.docx`, which variable recording the month of admission?, which variable recording the month of discharge?

'moa' records the month of admission.
'mod' records the month of discharge.

8. Which month admitted the most number of patients? Which month admitted the most number of male patients?

```{r}
library(dplyr)

df %>% group_by(moa) %>% summarise(patient_count = n()) %>% arrange(-patient_count)
```

```{r}
df %>% filter(sex == 1) %>% group_by(moa) %>% summarise(patient_count = n()) %>% arrange(-patient_count)
```

9. Which month has the most number of teenage female patients?

```{r}
df$age2 <- case_when(
          df$age < 13 ~ 'child',
          df$age < 20 ~ 'teenager',
          TRUE ~ 'adult')

df %>% filter(age2 == 'teenager', sex == 2) %>% group_by(moa) %>% summarise(patient_count = n()) %>% arrange(-patient_count)
```

10. Which provider has the most number of female patients in October? 

```{r}
df %>% filter(moa == 10, sex == 2) %>% group_by(provider) %>% summarise(patient_count = n()) %>% arrange(-patient_count)
```

11. Is female patients older than male patients, on average? 

```{r}
df_female <- df %>% filter(sex == 2)
df_male <- df%>% filter(sex == 1)

mean(df_female$age) > mean(df_male$age)
```

12. Calculate the average age of patients by months. Which month has the oldest patients on average age?

```{r}
df %>% group_by(moa) %>% summarise(average_age = mean(age)) %>% arrange(-average_age)
```

13. What is the name of the provider that has the highest total charge?

```{r}
df %>% group_by(provider) %>% summarise(total_charge = sum(tot)) %>% arrange(-total_charge)
```

Provider 7205 is Rhode Island Hospital.

14. What is the name of the provider that has the least total charge for teenage male?

```{r}
df %>% filter(sex == 1, age2 == 'teenager') %>%  group_by(provider) %>% summarise(total_charge = sum(tot)) %>% arrange(total_charge)
```

Provider 7214 is Woman and Infants.

15. Calculate the length of stays by races.  Which race has the longest length of stays on average?

```{r}
df %>% group_by(raceethn) %>% summarise(average_stay = mean(los)) %>% arrange(-average_stay)
```

Race 9 is unknown, so the longest stay by known race would be race 4 which is American Indian.

16. On average, how much a 20 year-old male white get charged for staying 1 day?

```{r}
df %>% filter(age == 20, sex == 1, raceethn == 1, los ==1) %>% summarise(average_charge = mean(tot))
```

-------

## II. Data Visualization

Continue with the data from part I. 

1. Provides at least 10 meaningful plots. Comments on the plots. All plots should have title, caption, appropriate labels on x and y-axis

```{r}
df %>% filter(sex == 1 | sex == 2) %>% 
ggplot(mapping=aes(x=age, y=tot, color=sex))+ geom_point()+ facet_wrap(~sex) +labs(title = 'Total Cost by Age', y = 'Total Cost', x = 'Age', caption = 'There is no significant difference between male and female for total cost by age.')
```

```{r}
df %>% filter(sex == 1 | sex == 2) %>% ggplot() + geom_bar(mapping=aes(x=raceethn, fill=sex), position='dodge') + labs(title = 'Count of Race by Sex', x = 'Race', y = 'Number of Patients', caption = 'All races had slightly more female patients than male patients.')
```


```{r}
df %>% filter(sex == 1 | sex == 2) %>% group_by(sex) %>% summarise(average_price = mean(tot)) %>% ggplot() + geom_col(mapping=aes(x=sex, y=average_price, fill=sex)) + labs(title = 'Average Price by Sex', x = 'Sex', y='Average Price', caption = 'Average price for male was about $5,000 higher than that for a female.')
```


```{r}
df %>% group_by(raceethn) %>% summarise(average_price = mean(tot)) %>% ggplot() + geom_col(mapping=aes(x=raceethn, y=average_price, fill=raceethn))+labs(title= 'Average Price by Race', x = 'Race', y = 'Average Price', caption = 'The highest average price was for race 1 (White) and the lowest average price was for race 3 (Asian).')
```

```{r}
df %>% group_by(age) %>% summarise(average_price = mean(tot)) %>% ggplot() + geom_col(mapping=aes(x=age, y=average_price, fill=age)) + labs(title = 'Average Price by Age', x = 'Age', y = 'Average Price', caption = 'This shows a large spike in average price around the ages 10 to 20.')
```

```{r}
df %>% group_by(age2) %>% summarise(average_price = mean(tot)) %>% ggplot() + geom_col(mapping=aes(x=age2, y=average_price, fill=age2)) +labs(title = 'Average Price by Age', x = 'Age', y = 'Average Price', caption = 'The average price for a teenager is the greatest.')
```

```{r}
df %>% filter(sex == 1 | sex == 2) %>% 
ggplot(mapping=aes(x=admtype, fill=sex))+ geom_bar()+ facet_wrap(~sex) +labs(title = 'Count of Admission Type by Sex', x = 'Admission Type', y = 'Number of Patients', caption = 'Admission Type 1 (Emergency) is the most common admission type for both males and females.')
```

```{r}
df %>% ggplot(mapping=aes(x=admtype, fill=age2))+ geom_bar()+ facet_wrap(~age2) + labs(title = 'Count of Admission Type by Age', x = 'Admission Type', y = 'Number of Patients', caption = 'All age groups have Admission Types 1 and 2 (Emergency and Urgent) in their most popular types.')
```

```{r}
df %>% filter(age2 == 'child' | age2 == 'teenager') %>% ggplot(mapping=aes(x=admtype, fill=age2))+ geom_bar()+ facet_wrap(~age2) + labs(title = 'Count of Admission Type by Age', x = 'Admission Type', y = 'Number of Patients', caption = 'Both child and teenager results resemble that of an adult. The exception is the large amount of admission type 4 (newborn) for child')
```

```{r}
df %>% group_by(age2) %>% summarise(average_los = mean(los)) %>% ggplot(mapping=aes(x=age2, y=average_los, fill=age2))+ geom_col() + labs(title = 'Average Length of Stay by Age', x = 'Age', y = 'Average Length of Stay', caption = 'The average length of stay for a teenager is greater than that of a child or an adult.')
```

2. Make an animation

```{r}
library(gganimate)
library(gifski)

df %>% 
  group_by(moa,age2) %>%
  summarise(average_los = mean(los)) %>%
  ggplot(aes(y=average_los,
            x=moa,
            color=age2,
            fill=age2))+ labs(title = 'Average Length of Stay by Age by Month of Admission', x = 'Month of Admission', y = 'Average Length of Stay')+
  geom_point(size=3)+
  geom_text(aes(label = average_los), 
            hjust = -.1, size=3)+
  geom_line()+
  transition_reveal(moa)
```

-------

## III. Predictive Models

Continue with the data from part I. Use the follows as the target and input variables: 

*Target Variable*: Create the target variable taking value of 

  - `low cost` if the total charge of a patient (`tot`) is smaller than the median of the total charge, and

  - `high cost` otherwise. 
  
```{r}
df$cost <- case_when(
          df$tot < mean(df$tot) ~ 'low_cost',
          TRUE ~ 'high_cost')
```

*Input Variables*:

  - "age","sex","raceethn","provider","moa","mod","admtype","campus", 'los'
 
```{r}
df1 <- select(df, age, sex, raceethn, provider, moa, mod, admtype, campus, los, cost)

df1 <- df1 %>% filter(raceethn != '', admtype != '')

```
  
-------

1. Make sure all the categorical variables are factor, numeric variables are numeric. Set Training : Testing Split = 10 : 90 

```{r}
library(mlbench)
library(caret)
set.seed(2020)

names(df1)[10] <- 'target'
splitIndex <- createDataPartition(df1$target, p = .90,
                                  list = FALSE)
df_train <- df1[splitIndex,]
df_test <- df1[-splitIndex,]
```

2. Train a decision tree using `rpart`.  Plot the decision tree. Plot the variable importance ranked by the tree. 

```{r}
library(rpart)
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))

library(rattle)
fancyRpartPlot(tree_model)
```

```{r}
barplot(tree_model$variable.importance)
```


3. Using caret for this question. Set `Training Control` to be: Use Cross-Validation of 5 folds across all models.  Train & tune at least 3 different models (i.e. three different values for `method=` in the train function of caret).  Plot the hyper-parameter tuning plots for each model. 

```{r}
tuneGrid = expand.grid(maxdepth = 1:5)

trControl = trainControl(method = "cv",
                         number = 5)

tree_approach1 <- train(target~., data=df_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach1)
```

```{r}
tuneGrid = expand.grid(maxdepth = 2:4, mstop=3)
tree_approach2 <- train(target~., data=df_train, 
                                method = "blackboost", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach2)
```

```{r}
tuneGrid = expand.grid(maxdepth = 2:4, iter=3, nu=2)

tree_approach3 <- train(target~., data=df_train, 
                                method = "ada", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)

plot(tree_approach3)
```

4. Plot the comparison of the models in 2. 

```{r}
results <- resamples(list(rpart2 = tree_approach1,
                          blackboost = tree_approach2,
                          ada = tree_approach3))
bwplot(results)
```

5. What is your final selection for the model? Test the accuracy of your final model on the test data. 

```{r}
pred <- predict(tree_approach3, df_test)
cm <- confusionMatrix(data = pred, reference = 
as.factor(df_test$target), positive = "high_cost")
cm$overall[1]
```


6. Create another `target` variable (binary), decide the input variables and redo 1 to 5. 

-------
```{r}
df2 <- select(df, age, sex, raceethn, provider, moa, mod, admtype, campus, los, cost)

df2 <- df2 %>% filter(raceethn != '', admtype != '', sex == 1 | sex == 2)
```

```{r}
names(df2)[2] <- 'target'
splitIndex <- createDataPartition(df2$target, p = .90,
                                  list = FALSE)
df2_train <- df2[splitIndex,]
df2_test <- df2[-splitIndex,]
```


```{r}
tuneGrid = expand.grid(maxdepth = 1:5)

trControl = trainControl(method = "cv",
                         number = 5)

tree_approach1_2 <- train(target~., data=df2_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach1_2)
```


```{r}
tuneGrid = expand.grid(maxdepth = 2:4, mstop=3)
tree_approach2_2 <- train(target~., data=df2_train, 
                                method = "blackboost", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach2_2)
```


```{r}
tuneGrid = expand.grid(maxdepth = 2:4, iter=3, nu=2)

tree_approach3_2 <- train(target~., data=df2_train, 
                                method = "ada", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)

plot(tree_approach3_2)
```


```{r}
results <- resamples(list(rpart2 = tree_approach1_2,
                          blackboost = tree_approach2_2,
                          ada = tree_approach3_2))
bwplot(results)
```


```{r}
pred <- predict(tree_approach3_2, df2_test)
cm <- confusionMatrix(data = pred, reference = 
as.factor(df2_test$target), positive = "1")
cm$overall[1]
```
